### 2016-12-04 treelapse.Rmd
Main report doc

---
title: "Interactive Visualization of Hierarchically Structured Data"
author: "Kris Sankaran"
date: "`r Sys.Date()`"
output:
	bookdown::pdf_book:
bibliography: ["bibliography.bib"]
link-citations: yes
---

```{r set-chunk-opts}
knitr::opts_chunk$set(fig.align = "center", out.width = "200px")
```

## outline ##

This is an outline for the treelapse article I want to write.

Introduction

- Give core theses
    + Using principles from the data visualization literature, we have designed
      a tool that we hope will be useful for analyzing tree structured data,
      especially hierarchically structured collections of time series.
    + Focus / context and linking are central to each of our visualizations
    + We hope our implementation makes this specific tool more widely
      accessible
- Describe paper structure

Problem Motivation

- Microbiome motivation
      + The meeting of ideas from medicine and ecology
      + The differential abundance problem: treatment vs. control in
        experiments
      + The ecological dynamics problem
             * Are there niches?
             * Is there competition?
- Abstraction
      + Random variables at nodes
	  + In differential abundance, we have a vector of counts, one coordinate
        per group
	  + For microbial dynamics, we have a time series at nodes
	  + In these applications, we define values at nodes as linear combinations
        of descendants (sum or average). This sort of constraint is not
		necessary though
- Problems with related structure
      + Spatial hierarchy when working with time series
      + Creating trees
             * Hierarchical clustering
             * CART
             * When variables don't have intrinsic order, but have comparable
               ranges, can interpret time series as parallel coordinates

Solution Principles / Background Literature

- Focus and Context (summary then examples)
      + Degree of interest -- a way of navigating graph structures. Modularize
        the problem of graph layout and interaction. One part is determination
        of a degree of interest for nodes. The other is determining a layout
        given a particular degree of interest distribution. Blocks are trimmed
        away depending on their average degree of interest.
      + Timeboxes -- A way of querying the types of shapes that appear in a
        given collection of time series. You draw boxes, and it highlights
        series that pass through them. The focus is on highlighted series, the
        context is the collection of series in the background.
      + Visual scent -- Determine pan and zoom based on a smaller version of
        the larger figure. Similar to interactively adjusting bounds
        numerically, but more intuitive. Focus is on the part of the scent that
        is highlighted, context is the rest of the scent.
- Linking (summary then examples)
      + Linked scatterplot brushes: Let you see many dimensions for the same
        few points. Conditional on having high values for one series, what are
        values for other variables? This is analogous to timeboxes on parallel
        coordinates plots.
      + Visual queries / conditional interpretation: Power is in querying one
        variable and seeing values in the other.
      + XGobi: Historical example of linking views, grounded in
        multidimensional scaling

- Participation in broader R software ecosystem
    + Can iterate through many interactive visualizations (like in ggvis)
    + Limited scope, but easier to maintain and more likely to remain relevant
    + Rely heavily on htmlwidgets
         * Also note similarity with networkD3: though, we have focused more on
           principles for a specific problem setup, rather than constructing
           foundational building blocks (in a way, this is more applied)

Specific Proposals

- DOI Trees
     + definition and (microbiome) use case: for a single sample, where are
       most bacteria?
     + principles it relies on: just DOI tree.
     + This is the only visualization that is completely thought through from
       before (but which as far as we can tell hasn't had easily accessible
       implementations).
     + well not exactly...we are also encoding abundance in edge widths / node
       sizes
- DOI Sankeys
     + Definition and use case: treatment vs. control in tree structured data,
       but using DOI principle
     + Exploratory analog of structSSI
     + Minor note: we use the colorbrewer palette for simpler readability
- Timebox Trees
     + Combines a few more ideas: linking and focus + context
     + definition and explanation of how these ideas being implemented
              * timebox recap: we're able to highlight lines passing through
                intersection of boxes (this is focus and context)
              * Link the display between tree and time series. Interpretation:
                Position on tree conditional on time series
                shape. Alternatively, database query on one field, returning
                over all of them.
              * Focus and context / pan and zoom using visual scent [so we have
                two focus and context ideas, and one linking idea, in this one
                display]
      + Primary use case this is designed for -- we notice a pattern in how a
        clump of microbial series look like, and we try to figure out which
        subtrees have this behavior (or, if that behavior is appear at
        different places on the tree).
- Treeboxes
       + Describe as the natural analog of timebox trees
       + Intentionally avoided queries on both tree and time series
       + Example use case: we can see how different parts of the tree have
         different series.
               * Or, we can look at all the phyla / orders at once
               * Or, we can observe if there is any competition within a
                 particular clade
       + Conditional on coming from a part of a tree, what are the observed
         series?

Case Studies

- Introduction
      + Practically, want to provide example output and how to interpret it
      + This is only a partial goal (there are many vignettes that describe
        this in more detail and are explicitly designed as tutorials) -- more
        of the focus here is on explaining how the visual principles described
        above can substantively improve data analysis in this problem domain
      + Purpose is to motivate potential workflows that incorporate the
        treelapse package. Show how it can be incorporated into any existing R
        data analysis workflow (this was all done in a single R document)
      + Argue that the problem of navigating hierarchical collections of time
        series is actually more generic than you might think

- Antibiotics
      + Study background: (check out abstract to confirm) How does microbiota
        respond to a time course of antibiotics. Especially relevant given
        growing understanding of importance of microbes in *health* (not just
        disease)
      + Framing as a timebox / treebox problem
      + Initial view: can clearly discern decrease across all microbes
      + Selecting some series, notice that the Firmicutes are more strongly
        affected
      + Can zoom into period around perturbation, using the scent idea
      + Can rerun using node averaged data -- this is only possible because we
        have linked it with R (it would not be possible in a standard d3 Block,
        for example)
      + Ideally: comparison with gPCA? Show how it complements.
- DOI Sankey: Preterm births
      + Basic goal of the study -- are certain microbiome signatures associated
        with preterm births?
      + Formal inferential approach was taken by the original publication,
        testing each microbe for an association with term / preterm
      + Also had an unsupervised analysis: cluster into CSTs
                - We can follow-up this unsupervised work using treelapse (this
                  is analogous to gPCA followup)
                - They used a heatmap, which gets the main points across, but
                  without the tree context (or allowing interactive focus +
                  context)
      + We immediately note that CST 4 has much more diversity
      + navigating down the tree, we notice that the main way the other CSTs
        are different is in the different types of Lactobacillus
- Housing prices
      + A non-microbial example, but with relatively obvious hierarchical
        (spatial) structure
      + Basic goal: How were home prices affected by recession, and how do
        prices and trends in prices differ across regions?
      + Findings from timebox / treebox:
                 * Can see quickly which neighborhoods are richer than others
                 * Can also see that poorer areas were more strongly affected
                   by the recession
                 * Can also see the variation in home prices within individual
                   counties
      + Limitations: We never actually see the spatial locations for the
        series. Can easily imagine a linking between the map and the original
        series (cite xgobi), but this is out of scope (at least for now)
- Bikesharing
      + A non-microbial example, and with non-obvious hierarchical structure
      + To induce hierarchical structure, we perform classification using a
        supplementary data table
                * We take advantage of the fact that this is essentially a two
                  table problem (one whose rows are different time series,
                  another giving the supplemental data)
                * Perform classification using classification trees (similar to
                  Kaggle, but not quite -- we treat a whole series as a sample,
                  not just a single time point).
       + Observations
                * Quickly notice difference between weekend and weekday
                * Also more subtle differences though -- temperature and year
                  effects, for example
                * The 3AM people on warm weekends (importance of focus)
       + Ideal scenario: Develop model using ideas from here, relate model to
         hierarchical structure again. Alternatively, fit a tree on correct and
         incorrect cases from another algorithm, inspect what are features of
         incorrectly classified cases, build these into new features.
- Potentially: genetics and hierarchical clustering?
- Summary: Tree structured data and collections of related time series are
  common than you might suspect, the principles of data visualization actually
  facilitate navigation of high dimensional data, and since it's in R, it's
  easy to alternate between interactive visualization, data regeneration /
  preparation, and statistical modeling.

Conclusion

- Potential future work
   + Arrangement
        * Cognostics for arranging views
        * Faceting timebox / treebox views
        * Linking with DPCOA / gPCA? In practice, microbiome literature cares a
          lot about these ordinations, we can try to explicitly tie them
          together in one visualization (rather than iterating, like we do
          now).
   + Incorporating the DOI idea into the timebox trees?
   + Other layouts? Graphs? DAGs?
   + Grammar
        * Is it possible to modularize the key ideas, so we can create similar
          kinds of linked and arranged views without so much effort?
- Summary
    + Developed an R package based on ideas from the visualization community
    + Think there is power in interactive visualization for hands on data
      analysis, not just polished public tools
    + Think there is a relationship between statistical modeling and data
      visualization: it can be more useful to visualize raw data in relation to
      models
    + Focused on solving a very specific methodological problem, rather than
      building a generic system. But this methodological problem appears in a
      variety of domains.

## text ##

## Introduction ##

In this work, we describe methods for visualization of data structured along
trees, especially hierarchically structured collections of time
series. Influenced by Degree of Interest trees, linked scatterplot brushing,
and timeboxes, we have adapted the fundamental visualization principles of
"focus + context" and "linking" to study questions that arise when analyzing
tree structured data. Our motivating application is to the analysis of
microbial time series, where an evolutionary tree relating microbes is
available a priori. However, we have identified common problem types where, if
a tree is not directly available, it can be constructed from data and then
studied using our techniques.

We have implemented our visualizations in d3, but encapsulated in an R package,
to facilitate rapid turnover from data preparation and modeling to interactive
exploration, and vice versa. We hope this package encourages data analysts to
work at this border of data modeling and visualization, and more generally
empowers a wider audience to apply less widely known, but powerful,
visualization ideas.

The paper is organized as follows. First, we describe our microbiome
application, and then abstract away the essential analysis tasks. This
abstraction suggests applications to standard problems outside of the
microbiome, some of which we articulate here. Next, we review the underlying
visualization principles behind our work, and provide references to the
software systems that originally implemented them. We then connect these
principles to analysis tasks we identified earlier, describing in detail the
visualization methods we have implemented in treelapse. We close with several
case studies using publicly available data.

### Problem Motivation ###

Microbiomics is the study of the communities of microbes living in given
environments, for example, ocean water or the human gut. Progress in the field
has been rapidly accelerated by the advent of genomic technologies, which
enable detailed quantification of microbial community structure, and as the
relationship between the microbiome and human health has become more and more
apparent, research into the human microbiome has garnered widespread
attention. Being concerned with both microbial community structure and human
health, the field exists at the border between ecology and medicine;
consequently, papers in the area often apply a blend of exploratory data
analysis and formal statistical inference ideas.

The two essential microbiome analysis problems that motivated our work are the
tree structured differential abundance and microbial dynamics problems. In the
differential abundance problem, we attempt to compare the abundances of
individual microbes across experimental conditions -- for example, treatment
vs. control or healthy vs. diseased. This is the microbiome analog of
differential expression analysis in genomics. We prepend the description
"tree-structured" because, in practice, researchers generate interpretations
about middling taxonomic orders of microbes -- it is more interesting to
discover novel behavior taxonomic levels intermediate between high-order phyla
and low-level sequence variants. Hence, we frame the tree-structured
differential abundance problem as the question of identifying the largest
taxonomic subtree whose associated microbes are differentially abundant.

In the tree structured microbial dynamics problem, the goal is to describe
changes in microbial abundances in an environment over time. As in the
differential abundance problem, it is useful if these descriptions can be given
at the highest subtree at which the pattern appears. Specific questions of
interest often have an ecological flavor. For example, researchers are often
interested in understanding how microbial populations respond to sudden or
gradual environmental changes or how microbes fill, drop out from, or compete
for certain environmental niches. Medically, these questions are important for
illuminating the impact of antibiotic time courses or the potential for direct
manipulation, for example.

### Problem Abstraction ###

An abstraction that unifies thinking through both the tree structured
differential abundance and microbial dynamic problems is to identify the data
with a collection of random variables indexed by nodes in a prespecified tree
structure. In the differential abundance problem, each random variable itself
lives in $\mathbb{R}^{G}$ where $G$ is the number of groups being
compared. Each coordinate represents the average value across one of the
groups, and the node exhibits differential abundance when the coordinates are
not drawn from the same distribution. In the microbial dynamics problem, each
random variable is instead a time series, living in $\mathbb{R}^{T}$.

In both these applications, we constrain the values of parent nodes according
to the value of the children nodes: we define the value at each node to be
either the sum or average of all descendant tips, but it is possible to imagine
situations where the internal nodes are drawn from their own distribution,
unconstrained by descendants. In general, analysis in this abstraction focuses
on describing the distribution of these random variables as a function of their
position in specified subsets of the tree. The essential difficulty in these
problems is high-dimensionality -- there are many tree nodes, each holding a
vector-valued random variable. Even simply navigating across the tree and
comparing coordinates in the observed variables is a challenge; ideally we
could construct a succinct representation of the essential covariation across
subtrees and coordinates.

This framework suggests other potential application areas, not all of which
have a priori known tree structures. For example, collections of spatially
indexed time series are frequently encountered in practice -- consider energy
consumption, product sales, or high school dropout rates across regional
districts. This type of data has an implicit tree structure -- at the top level
are different states, while at the bottom are individual census tracts,
say. Analysis here revolves around the question of how variation across series
is related to their geographic position.  An example using this type of data is
given in the Zillow home prices case study in Section (todo: insert section
number).

Alternatively, if this type of hierarchical contextual information is not
available, a tree structure can be learned directly from the data. This can be
most directly achieved by learning a hierarchical clustering on the original
series. If there is contextual information, but it is not hierarchical, it is
possible to setup a supervised problem that uses context to predict features of
the time series. We can retrieve a tree by using a regression tree directly
(cite: rpart) or extracting a regression tree from a more complex supervised
model (cite: regression tree extraction methods). Analysis then focuses on how
different partitions of the covariate (contextual) space relate to observed
time series. This approach is described in the UCI bikesharing demand case
study below.

Finally, note that, while we have focused on time series valued nodes, all of
this discussion could be translated to studying high dimensional data via
parallel coordinates (todo: cite parallel coordinates). The usual parallel
coordinates challenges remain, mainly selecting scales for and ordering the
different coordinates, but the linking and focus + context can still be
employed this setting.

## Background Literature and Solution Principles ##

Now that we have specified the essential questions of interest, we survey some
ideas from the visualization literature that can be applied to answer
them. Again, the core difficulty is one of high dimensionality, so it should be
no surprise that the techniques we adapt come from the literature on
high-dimensional data visualization, namely, work on the focus + context and
linking.

The focus + context principle is that it is possible to study large collections
of visual elements at multiple scales, and these scales can be juxtaposed by
"focusing" on a few elements of interest while maintaining the "context" of the
coarser-scale background. A simple example of this idea is to include a search
box that highlights samples corresponding to matching terms (the focus) while
muting those that don't (the context). Two more sophisticated methods anchored
in this idea are timeboxes and Degree-of-Interest (DOI) trees; both are central
to the proposals in treelapse \@cite(hochheiser2004dynamic,
heer2004doitrees). In timeboxes, a collection of time series are graphically
queried using interactive brushes. Series that pass through all of the
user-specified brushes are highlighted, and the rest are faded into the
background. Hence, time series meeting the constraints imposed by the brushes
are focused, while the remainder remain present, though de-emphasized, as
context. This method can be interpreted programmatically as the visual version
of a database query, or probabilistically as the conditional distribution for
the full series, given it passes through certain bounds.

In DOI Trees, the viewer's attention is focused on a collection of
high-interest nodes, while the remaining lower-interest nodes are left on the
fringes (or entirely hidden) as context. The implementation is modularized into
two tasks -- the determination of a DOI distribution over nodes in the tree and
visual layout of a tree given DOI assignments. The DOI distribution used in
\@cite(heer2004doitrees) places maximal interest on nodes that the user clicks,
along with all ancestors. The DOI for all other nodes is defined as the
distance to the closest maximal interest node. The layout proposed here trimmed
low-interest subtrees until the remaining nodes all fit within the screen
size. By changing the minimal DOI below which nodes are hidden, the user can
transition between node-specific and full-tree scales.

In linking, alternative representations of the same samples are placed side by
side in a way that suggests covariation across the representations. A canonical
application is to linked scatterplot brushing \@cite(becker1987brushing). Here,
a scatterplot matrix is constructed displaying the relationship between all
pairs of variables.  Points brushed in one scatterplot are then highlighted in
all others. For example, this helps the user determine whether an outlier in
one dimension is an outlier in others. Another instance of this idea is in
linking the results of dimensionality reduction methods to displays of the raw
data, as implemented by XGobi and Cranvas, for example. As in timeboxes,
linking ca ben interpreted as database queries or conditional probabilities:
given a subset of the series after conditioning on the values for one set of
features, what are the values for a second set \@cite(buja1996interactive)?

Finally, while not exactly a fundamental visualization principle, we note that
our work is deliberately grounded in the R software ecosystem. Not only does
this make these methods more broadly accessible, we hope to encourage a more
animated exchange between data preparation and modeling with interactive
visualization. Moreover, our tools are intentionally limited in scope --
designed to facilitate this dialog for a specific class of problems, rather
than providing a toolbox for generic types of visualization design. We believe
that this narrow context within a broad ecosystem strikes a balance between
problem-specificity and ease-of-use.

## Specific Proposals ##

Our first proposed visualization technique is a minor modification of the DOI
Tree idea. The standard DOI Tree definition does not have any notion of data
defined at nodes, it is only used a device for navigating tree structures. A
trivial extension can encode scalar data at nodes: make the node radius reflect
the associated scalar value. To reinforce this effect, we also encode this
scalar value in the width of the parent edge. When parent nodes have values
equal to the sum of their children, this creates the effect of values "flowing"
from the root to leaves. To help viewers make use of their domain knowledge, we
have included a search box that highlights paths to matching node terms. Edges
are ordered from widest on the left to narrowest on the right. While this
method can only be used to visualize a single scalar-value per tree node, it
suggests an approach to the tree-structured differential abundance problem,
which we call the DOI Sankey.

In the DOI Sankey, we split each edge in the DOI Tree across different
groups. For example, suppose we have the average counts for treatment and
control groups at each tree node. Every edge in the tree is split into two
colors^[We use the colorbrewer palette to facilitate readability], with the
relative widths of the different colors reflecting the difference in sizes for
the two groups. The overall width of each edge is represents the sum of values
across all groups.

This display is designed to facilitate investigation of the tree structured
differential abundance question. For example, for a single node and a single
group, we can compute the average abundance at that node among all samples in
that group. This will give the width for that group's color on the edge leading
to the specified node. Differentially abundant subtrees then correspond to
subtrees where some colors occupy more space than others. That is, this
representation makes it easier to identify points where the "flows" for
different groups diverge -- the colors begin to separate. The DOI principle
assists navigation across the tree structure, allowing focus on individual flow
structures without losing broader tree context.

Our third display is directed at the microbial dynamics question instead. It
ignores the DOI principle, placing emphasis rather on linking and the
representation of full time series. Here, two panels are arranged one over the
other; one displays all time series, while the other displays all tree nodes,
with node sizes reflecting the average time series value at that node. For this
reason, we call the display, Timebox Trees. In the time series panel, we have
directly implemented the timeboxes idea. Our link is between the two panels:
when a set of series is highlighted by the timeboxes, the associated tree nodes
are also highlighted. For example, timeboxes can be used to focus on a set of
series with specific shape -- increased abundance after an ecological shock,
for example -- and then identify along what subtrees this pattern is
present. To further focus attention, a pan-zoom scented widget is provided in
the top right [@willett2007scented]. The widget is a miniature version of
the full time series panel, equipped with a single brush whose extent specifies
the limits in the main time series panel. As in the DOI trees and Sankeys, a
search bar can be used to highlight of interest a priori.

The final display currently implemented in the package is the natural converse
of the Timebox Trees display. Rather than defining visual queries in terms of
time series, we define queries using nodes in the tree. For this reason, we
call the display Treeboxes. Rather than focusing on the intersection of
brushes, as in Timebox Trees, we focus on the union of brushed over nodes. This
allows us to highlight the series associated with nodes on distant
subtrees. This display is also suited for the microbial dynamics problem. For
example, by highlighting all nodes at one taxonomic level in the tree, we can
easily summarize the time series pattern for all the taxa at that
level. Alternatively, focusing on all the children below a single node makes it
possible to see how much correlation and competition there is between
taxonomically similar microbes. As in the Timebox Trees display, a search box
and pan-zoom scented widget is provided.

## Case Studies ##

We now delve into concrete examples of applying treelapse to real
data. Our goals are is to illustrate potential workflows that
incorporate treelapse, describe the formulation of questions that can
be naturally investigated with our methods, and provide example
interpretations for specific treelapse output. Our examples are also
chosen to reflect the range of problem domains to which the package
can be applied -- though it was motivated by applications to the
microbiome, it is not tied to it. More importantly, we argue that the
visualization principles described earlier can substantively improve
the practice of data analysis in the class of problems we have focused
attention on.

### Microbial Dynamics of Antibiotics Time Courses ###

[@dethlefsen2008pervasive] investigates the effect of antibiotics on microbial
community composition from an ecological perspective. The study tracks the
microbiome of *x* patients across *y* weeks (todo: give the numbers), with
antibiotic two time courses introduced during * and *, respectively. Discerning
the variation in resilience across microbes is important, considering the
community's increasing acceptance of the idea that microbes are central to human
health and not just disease.

We approach the data using the linked time and treebox views, after first
filtering low variance taxa and taking an $asinh$ transformation. An initial
view, Figure \@ref(fig:antibioticoverview), reveals two dramatic drops in the
overall microbial abundance time series at the time of the antibiotics time
courses. Two more subtle effects are also suggested from this view,

* The second antibiotic treatment seems to have a more lasting effect, as the
series take longer to return to their original values.
* Some high level taxa appear relatively unaffected by the first
antibiotic treatment.  By more closely inspecting the display, we are
able to identify these as members of the Bacteroidetes phylum, see
Figure \@ref(fig:antibioticbacteroidetes).

```{r antibioticoverview, fig.cap = "The two antibiotic time courses are readily apparent even when only highlighting the most abundant taxa."}
knitr::include_graphics("antibiotic_overview.png")
```

```{r antibioticbacteroidetes, fig.cap = "Introducing a second box into the timebox display identifies the Bacteroidaceae as a taxon only midly impacted by antibiotics."}
knitr::include_graphics("antibiotic_bacteroidetes.png")
```

Next, using the scented widget, we focus on the window around the
second antibiotic treatment. We apply the treebox display to compare
then behavior of different families of Firmicutes, Lachnospiraceae and
Ruminoccocus. We suspect these taxa are associated with the delayed
recovery after the second time course. To inspect this, we these
family names in the search box to isolate their positions on the tree,
and then we apply brushes to highlight the series that contribute to
these higher-level families. The resulting view is given by Figure
\@ref(fig:antibioticfirmicutes)

```{r antibioticfirmicutes, fig.cap = "Zooming into the second antibiotic timecourse and highlighting the Lachnospiraceae and Ruminococcus, we see that the Ruminoccocus took more time to recover to pre-treatment levels."}
knitr::include_graphics("antibiotic_firmicutes.png")
```

Alternatively, we can summarize each node by the average of its descendants --
this brings attention to individual microbes that may be underlying some of the
broader taxonomic patterns we have noted when studying the subtree sums. For
example, in Figure \@ref(fig:ruminococcus), we highlight all families below the
Ruminococcus, suggesting that the decrease due to antibiotics occurs uniformly
across almost all families. A point that was not evident in the earlier
sum-across-descendants view is that, after the second treatment of antibiotics,
a few of the Ruminoccocus families recover more rapidly than the rest, for
example the Unc095d3 (highlighted in red) are only briefly affected. In
contrast, most families seem to recover in unison after the first treatment.

```{r ruminococcus, fig.cap = "By hovering over the Ruminoccocus branches, we see that there is a prolonged effect of the antibiotics time courses more or less uniformly across the lower taxonomic orders, with the except of "}
knitr::include_graphics("ruminococcus.png")
```

Further, note that in this subtree averages view, the tree display has
changed. This is because, at each branching point, we place the node
with larger average value on the left. This places the
Verrucomicrobiae at the far left, which seem to have large abundance
over many time points. This phylum had been previously obfuscated,
because there are not many leaves associated with this phylum, hence
the sum was small. Interestingly, the abundance of these microbes
seems to *increase* after the first antibiotics treatment.  Be
cautious, however, that the average over only a few Verrucomicrobiae
species will be a more variable estimate than the averages over the
more prevalent phyla.

```{r verrucomicrobiae, fig.cap = "The subtree averages aggregation brings attention to the Verrucomicrobiae, which though only present as a few species, are each rather abundant. In particular, they seem to increase after the first antibiotic time course."}
knitr::include_graphics("verrucomicrobiae.png")
```

### Differential Microbial Abundance and Preterm Births ###

The study [@digiulio2015temporal] tracked the abundance of microbes in the
during vaginal microbiome during pregnancy in an effort to study relationships
between microbial community composition and preterm birth. Ideally, it would be
possible to develop clear microbial signatures associated with preterm births.

Unlike the antibiotics study, we have measurements across more individuals than
we could reasonably inspect one at a time. While we could average across all
individuals, we will take our cue from [@digiulio2015temporal] and place each
sample into one of five Community State Types (CSTs), identified via k-medoids.
Using a linear model, one of these CSTs (CST 4) was found to be more diverse,
further it appeared associated with preterm births. Here, we corroborate this
finding using exploratory views.

Therefore, our focus here is on the differential abundance question,
rather than dynamics.  We would like to provide visual representations
of differential abundance across CSTs and also between preterm and
non-preterm births. [@digiulio2015temporal] interpreted its CSTs using
a heatmap, with microbes ordered according to a hierarchical
clustering.  By using the DOI Sankey instead, we can interpret the
CSTs in their taxonomic context and at multiple scales of taxonomic
resolution. Further, while [@digiulio2015temporal] focused on
identifying associations between preterm births and CSTs -- presumably
because testing individual microbes loses power -- we can compare
microbe abundances between preterm and non-preterm samples along
subtrees, without requiring CSTs as an intermediary.

In Figure \@ref(fig:pretermcsts), we compare the 5 CSTs according to
their values along the subtree. Specifically, we took the average of
all samples within each CST to define values at the (microbe-level)
leaf nodes, and then aggregated the counts up to the root. It is
immediately clear that samples from CST 4 have much more taxonomic
diversity. Further, focusing on the Lactobacillaceae family, we note
that the differential abundnace of these microbes distinguishes th
remaining CSTs, see Figure \@ref(fig:pretermcstslacto).

Alternatively, in Figure \@ref(fig:preterm), we avoid working with
CSTs, displaying instead averages among samples associated with either
preterm or term births. The green edges are associated with preterm
births -- we see that they contribute more weight to Phyla outside the
Firmicutes. This is consistent with the claim that CST 4, the most
diverse of the CSTs, is associated with preterm births.

```{r pretermcsts, fig.cap = "The increased diversity among samples in CST 4 is represented by the relatively larger contribution of red edges to branches outside of the Firmicutes."}
knitr::include_graphics("preterm_csts.png")
```

```{r pretermcstslacto, fig.cap = "Zooming into the Lactobacillaceae family, we notice that the difference between the remaining four CSTs is related to which types of Lactobacillus are most prominent."}
knitr::include_graphics("preterm_csts_lacto.png")
```

```{r pretermpreterm, fig.cap = "Samples with high levels of phyla other than Firmicutes appear to be related to preterm births."}
knitr::include_graphics("preterm_preterm.png")
```

### Dynamics in Housing Prices ###

We next consider an application unrelated to the microbiome, but with
relatively clear hierarchical structure. Our data are downloaded from
Zillow^[http://www.zillow.com/research/data/](http://www.zillow.com/research/data/)],
and give the Zillow Home Value Indexes at the neighborhood level,
across the country, given monthly between 1996 and 2016. In our
display, we have taken the base 10 log of these indexes. As our
hierarchical structure, we use each neighborhood's assignment to
state, regional, county, and city levels. We represent each of these
coarser spatial categories using the average of all neighborhoods
contained in them. We have filtered down to the 890 neighborhoods in
California; rendering more neighborhoods while keeping all 246
timepoints causes the interface to lag^[See the Future Work section
for potential optimizations, however.].

Our basic analysis revolves around geographic and temporal variation
in home prices. We are especially interested in the effect of the 2008
recession and any variation in the lead-up or recovery. These
questions can be naturally framed using timebox trees and treeboxes.

We first study variation in prices at the city level. These series can
be highlighted using a single brush in treeboxes, since each spatial
level is displayed at the same height in the tree. THe resulting
display is given in Figure \@ref(fig:zillowcities). From a high level,
the housing bubble, decrease in prices due to the recession, and
subsequent recovery are readily apparent. The city-level view makes it
clear that not all neighborhoods were equally affected by the
recession -- richer cities plateaued at their 2008 prices,
middle-income cities saw moderate decreases, and poorer cities saw the
most significant declines in home prices. These more significant
declines tended to occur in cities that had recently seen rapid
increases during the housing bubble. Finally, we note that the range
in home prices seems to have increased after the recession, speaking
to the differential long-term effect of the recession on prices.

Alternatively, we can consider the trajectories of home prices among
neighborhoods, conditional on their being that were middle-income
before the recession.  We generate the sequence of views in Figures
\@ref(fig:zillowmiddlepre), \@ref(fig:zillowmiddleup), and
\@ref(fig:zillowmiddledown) in this vein. The first of these figures
isolates neighborhoods with middle incomes before the recession, using
a single timebox. Since there appears to be a divergence in
trajectories after the recession, we introduce a second post-recession
timebox, dragging it over series with higher and lower incomes during
this second time period. This is the content of Figures
\@ref(fig:zillowmiddleup) and \@ref(fig:zillowmiddledown). Inspecting
the highlighted tree nodes associated with these series, we find that
most of the middle-income series that increased after the recession
are associated with middle-income neighborhoods within the coastal
Southern California counties. In constrast, those middle income series
that saw decreases were mostly located in Central California and
Oakland.

The previous analysis highlights the fact that, within even narrow geographic
regions, there can be substantial variation in prices. We can study this
directly using treeboxes. In Figure \@ref(fig:zillowsf) we have highlighted all
series in San Francisco County. We see that, in 2016, prices range from around
$10^{5.6} \approx \$400,000$ to $10^{6.7} \approx \$5 \text{million}$. So, while
all these neighborhoods tend to be among the more expensive ones in California,
prices can vary in a non-smooth way across geographic space.

We conclude this example with a caveat that the Zillow data are not
representative of all neighborhoods in California, only those with
enough listings on the site, so should be supplemented by other data
sources for more substantial decision-making.

```{r zillowcities, fig.cap = "California home prices at the city level, between 1996 and 2016. The effect of the 2008 recession is clear, and we have hovered over the San Bernardino series, to display the identity of one of the cities most strongly affected by the recession."}
knitr::include_graphics("zillow_cities.png")
```

```{r zillowmiddlepre}
knitr::include_graphics("zillow_middle_pre.png")
```

```{r zillowmiddleup}
knitr::include_graphics("zillow_middle_up.png")
```

```{r zillowmiddledown}
knitr::include_graphics("zillow_middle_down.png")
```

```{r zillowsf}
knitr::include_graphics("zillow_sf.png")
```

### Sources of Variation in Bikesharing Demand ###

Our next example is a study in bikesharing demand, included as an
example of analyzing collections of time series when there is no
obvious hierarchical structure a priori. The data are downloaded from the
[UCI Repository](https://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset) and
were originally collected by a Washington D.C.-based bikesharing
system for use in a Kaggle prediction competition. The data are hourly
measurements of bike demand, aggregated across all bikesharing
stations, over two years, along with some supplemental weather
data. In the competition, participants were asked to predict the
hourly demand on a held-out test set. Here, we adopt a descriptive
view instead, attempting to characterize factors associated with
variation in bikesharing demand.

Like the Zillow home prices application, we study this problem as one
of describing a large collection of related time series. Here, we
consider the demand during a single day to be one time series; this is
a natural choice considering the daily periodicity of bike demand. To
relate these daily series according to an interpretable tree
structure, we apply a regression tree relating the supplemental data
to the bikesharing demand (J.  Friedman, Hastie, and Tibshirani
2001). In more detail, we build this tree by noting the “two table”
structure of this problem: one describes bike demand, the other holds
the supplemental data. In both, the rows index days, while the columns
index hours and supplemental features for the two tables,
respectively. Our tree is the trained regression tree after predicting
demand at 8AM based on the supplemental data. We choose this response
because (1) we need a univariate response in order to apply regression
trees and (2) the more straightforwards reduction to
daily-average-demand fails to distinguish between weekdays and
weekends, whose series appear qualitatively very different from each
other.

Given this response, the first split in the regression tree is
(unsurprisingly) the difference between weekends and weekdays. This is
emphasized in Figures 14 and 15, respectively; using timeboxes to
isolate the two types of series highlight the left and right sides of
the tree, respectively. For a mroe subtle effect, we select the
internal nodes associated with the first split below the weekday
vs. weekend split; these are given in Figures 16 and 17. This suggests
that more commuters became aware or interested in bikesharing during
the second year of data collection.

To contrast with these general questions about daily demand, we ask a
more granular question about demand at night: what characterizes days
on which there is larger than average demand after midnight? We can
select these series after first zooming into this time window. Figure
18 reveals that the highlighted series are associated with the
warm-weekend split, which seems quite reasonable in retrospect.

Finally, we can study the behavior of the regression tree itself using
the DOI Sankey. Here, we group samples according to their quintile of
8AM demand and then count the abundance of the groups flowing down
different branches. We find that the quintiles are each rather
strongly separated after descending even a few steps down the
regression tree – for example, in Figure ?? focuses on 2011 vs. 2012
split among weekday samples, showing that this split distinguishes
between samples falling in the second and third quintiles of 8AM
demand.

This interactive representation of regression trees is potentially
more useful on larger trees that cannot be easily parsed in a single
view; in this sense the bikesharing tree is relatively
simple. Further, in light of research on approximating more complex
supervised learning algorithms using decision trees (Boz 2002, Saito
and Nakano (2002)), this approach of modeling and interpreting
two-table data could be used in conjunction with more sophisticated
regression methods. In our ideas data analysis workflow, we imagine
the analyst applying interactive visualization and modeling techniques
in an iterative, nonlinear fashion.

## Conclusion and Future Work ##

Here, we have reviewed some fundamental principles of data visualization and described their implementation
in a new treelapse package. Further, we have provided examples of the practical usefulness of these
principles in real-world data analysis situations.

This package has only developed some of the most basic ideas, and
there are a number of potentially useful extensions worth
exploring. For example, we have not considered the principle of
arrangement in our visualizations (Buja, Cook, and Swayne 1996),
though many of our conclusions were based on comparing alternative
selections of the same display. We could imagine faceting our displays
across groups to make these types of comparisons more
accessible. Further, we have only worked with the DOI distribution
described in (Heer and Card 2004). It would be interesting to define a
more statistical notion of interest along nodes, based on cognostics,
for example (Hafen et al. 2013, J. H. Friedman and Stuetzle
(2002)). Further, it could be helpful to link the visualization-based
workflow described in this work with more traditional multivariate and
inferential statistics workflows – it is easy to imagine an
interactive version of ordinations that display representations of the
raw data upon selection. A simple extension could be to allow graph
layouts instead of trees in time and treebox displays, for data that
cannot be coerced into a hierarchical structure. Finally, if these
ideas turn out to be useful in practice, it would be valuable to
modularize the basic visualization layouts and relationships into a
library, allowing the wider community to construct linked, interactive
graphics with minimal effort.

All these potential extensions aside, we have built an easily
accessible R package for visualization techniques in a very specific
methodology problem – analysis of differential abundance and dynamics
in hierarchical structured data – that appears in a variety of
application domains. We have leveraged (Vaidyanathan et al.  2014) to
create exploratory visualizations while still deep in data analysis;
in this way our work is a departure from the culture of polished,
journalistic visualizations prioritized by the D3 community. Finally,
we have given a series of examples to demonstrate how the general
visualization techniques of focus + context and linked brushing can be
practically incorporated into a range of practical analysis workflows,
from studying the impact of microbes on human health to better
allocating units in commuter bikesharing systems.

## References ##