### 2017-07-12 treelapse.tex
Add description of contributions / context

\documentclass[12pt]{article}
\usepackage{hyperref}

\usepackage{amsmath, amssymb, amsfonts}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Interactive Visualization of Hierarchically Structured Data}
  \author{Kris Sankaran\thanks{
      This research was supported in part by an NIH training grant (5T32GM096982-03) and a Weiland Fellowship.}\hspace{.2cm}\\
    Department of Statistics, Stanford University\\
    and \\
    Susan Holmes\thanks{
       This research was supported in part by NIH grant R01 AI112401.
      }\hspace{0.2cm}\\
    Department of Statistics, Stanford University}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Title}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
We introduce methods for visualization of data structured along trees,
especially hierarchically structured collections of time series. To
this end, we identify questions that often emerge when working with
hierarchical data and provide an R package to simplify their
investigation. Our key contribution is the adaptation of the
visualization principles of focus-plus-context and linking to the
study of tree-structured data.

Our motivating application is to the analysis of bacterial time
series, where an evolutionary tree relating bacteria is available a
priori. However, we have identified common problem types where, if a
tree is not directly available, it can be constructed from data and
then studied using our techniques. We perform detailed case studies to
describe the alternative use cases, interpretations, and utility of
the proposed visualization methods.

\end{abstract}

\noindent%
{\it Keywords:}  D3, focus-plus-context, linking, time-series, tree-structured, R
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\section{Introduction}\label{introduction}

We introduce methods for visualization of data structured along trees,
especially hierarchically structured collections of time series. We hope both to
characterize generically useful techniques for interactively visualizing
hierarchical data and to offer practical tools for implementing such displays.
To this end, we identify questions that often emerge when working with
hierarchical data and provide an R package to simplify their investigation
\citep{ihaka1996r}.

In particular, we adapt the visualization principles of focus-plus-context and
linking to the study of tree-structured data \citep{buja1996interactive,
  becker1987brushing}. Our motivating application is to the analysis of
bacterial time series, where an evolutionary tree relating bacteria is available
a priori. However, we have identified common problem types where, if a tree is
not directly available, it can be constructed from data and then studied using
our techniques.

We have implemented our visualizations in D3, but encapsulated in an R package,
called treelapse, to facilitate rapid turnover from data preparation and
modeling to interactive exploration, and vice versa. Our code is open-source,
and available at \url{https://github.com/krisrs1128/treelapse}. We hope this
package encourages data analysts to work at the border between data modeling and
visualization, and more generally empowers a wider audience to apply less widely
known, but powerful, visualization ideas.

In summary, our key contributions are,
\begin{itemize}
\item Proposals for visualizing hierarchically structured data, based on
  principles from the data visualization community.
\item The implementation of these proposals in a publicly available R package.
\item The illustration of the wide reach of hierarchical data visualization,
  through case studies in both scientific and societal contexts.
\end{itemize}

The paper is organized as follows. First, we describe our motivating application
to the microbiome and the associated generic analysis tasks. Next, we review the
underlying visualization principles behind our contributions. Then we then
connect these principles to analysis tasks we identified earlier, describing in
detail the visualization methods we have implemented in treelapse. We close with
several case studies using publicly available data across both microbiome and
non-microbiome related applications.

\subsection{Problem Motivation}\label{problem-motivation}

A microbiome is a community of bacteria living in given environments, for
example, ocean water or the human gut \citep{human2012structure, cho2012human}.
Progress in the field has been rapidly accelerated by the advent of genomic
technologies, which enable detailed quantification of bacterial ecological
structure and its influence in human and environmental health. Being concerned
with both bacterial community structure and human health, the field exists at
the border between ecology and medicine; consequently, papers in the area often
apply a blend of exploratory data analysis and formal statistical inference.

The two essential microbiome analysis problems that motivated our work are the
tree-structured differential abundance and differential dynamics problems. In
the differential abundance problem, we attempt to compare the abundances of
individual bacteria across experimental conditions -- for example, treatment vs.
control or healthy vs. diseased. This is the microbiome analog of differential
expression analysis in genomics \citep{anders2010differential}. We prepend the
description ``tree-structured'' because, in practice, researchers generate
interpretations about intermediate taxonomic orders -- it is more interesting to
discover novel behavior taxonomic levels between high-order phyla and low-level
species. Hence, we frame the tree-structured differential abundance problem as
the question of identifying the largest taxonomic subtree whose associated
bacteria are differentially abundant.

In the tree-structured bacterial dynamics problem, the goal is to describe
changes in bacterial abundances in an environment over time. As in the
differential abundance problem, it is useful if these descriptions can be given
at the highest subtree at which the pattern appears. Specific questions of
interest often have an ecological flavor. For example, researchers are often
interested in understanding how bacterial populations respond to sudden or
gradual environmental changes or how species fill, drop out from, or compete for
environmental niches. Medically, these questions are important for illuminating
the impact of antibiotic time courses or diet changes, for example.

\subsection{Problem Abstraction}\label{problem-abstraction}

To unify the tree-structured differential abundance and bacterial dynamics
problems, we identify the data with a collection of random variables indexed by
nodes in a prespecified tree structure. In the differential abundance problem,
each random variable lives in $\mathbb{R}^{G}$ where $G$ is the number of
groups being compared. Each coordinate represents the abundance for that group,
and a node exhibits differential abundance when the coordinates are drawn from
different distributions. On the other hand, in the bacterial dynamics problem,
each random variable is a time series, living in $\mathbb{R}^{T}$.

In both of these applications, we constrain the values of parent nodes according
to the value of the children nodes: we define the value at each node to be
either the sum or average of all descendant tips. However, it is possible to
imagine situations where the internal nodes are drawn from their own
distribution, unconstrained by descendants. In general, analysis in this
abstraction focuses on describing the distribution of these random variables as
a function of their position across subsets of the tree. The essential
difficulty in these problems is high-dimensionality -- there are many tree
nodes, each holding a vector-valued random variable. Even simply navigating
across the tree and comparing coordinates in the observed variables is a
challenge; ideally we could construct a succinct representation of the essential
covariation across subtrees and coordinates.

This framework suggests other potential application areas, not all of which have
a priori known tree structures. For example, collections of spatially-indexed
time series are frequently encountered in practice -- consider energy
consumption, product sales, or high school dropout rates across regional
districts. This type of data has an implicit tree structure -- at the top level
are different states, while at the bottom are individual census tracts, say.
Analysis here revolves around the question of how variation across time series
is related to their geographic position.

Alternatively, if this type of hierarchical contextual information is not
directly available, a tree structure can be learned from the data. This could be
achieved by learning a hierarchical clustering on the original series. Further,
if contextual information is available, but it is not hierarchical, it is
possible to setup a supervised problem that uses context to predict features of
the time series. We can construct a tree by applying a tree-based classifier
\citep{breiman1984classification} or extracting a regression tree from a more
complex supervised model \citep{boz2002extracting,saito2002extracting}. Analysis
then focuses on how different partitions of the contextual, covariate space
relate to observed time series.

Finally, note that, while we have focused on time series valued nodes,
all of this discussion could be translated to studying high-dimensional
data via parallel coordinates \citep{inselberg1991parallel}. The usual parallel
coordinates challenges remain, mainly selecting scales for and ordering
across the different coordinates, but the linking and focus-plus-context
can still be employed this setting.

\subsection{Background Literature and Solution
Principles}\label{background-literature-and-solution-principles}

Now that we have specified the essential questions of interest, we survey some
ideas from the visualization literature that can be applied to answer them. As
the core difficulty is high-dimensionality, it should be no surprise that the
techniques we adapt come from the literature on high-dimensional data
visualization, which has enjoyed rapid progress in the last 25 years. Modern
research in this domain develops abstractions and taxonomies for guiding
visualization designs so that they most effectively communicates properties of
the data to their intended audience. A major push in this body of work explores
the potential for interactivity to improve many stages of the data analysis
process, from preliminary data preparation, to refinement and navigation across
views, to final sharing and annotation of results \citep{heer2012taxonomy}.
Further work has attempted to bridge the gap between statistical analysis and
data visualization methodology, both of which provide techniques for learning
from high-dimensional data \citep{de2003visual}.

The problem structures most relevant to our study are tree and temporal
structure, and the visualization community has various ways of reasoning about
these data, see \citep{graham2010survey, aigner2011visualization} for detailed
surveys. From this literature, our approach is most directly informed by the
focus-plus context and linking principles, which we briefly review here. The
focus-plus-context principle is that large collections of visual elements can be
studied at multiple scales, by simultaneously focusing" on a few elements of
interest and maintaining the ``context'' of the coarser-scale background. A
simple example of this idea is to include a search box that highlights matching
samples (focus) and mutes the rest (context). Two more sophisticated methods
anchored in this idea are timeboxes and Degree-of-Interest (DOI) trees; both are
central to the proposals in treelapse \citep{hochheiser2004dynamic,
  heer2004doitrees}. In timeboxes, a collection of time series are graphically
queried using interactive brushes. Series that pass through all of the
user-specified brushes are highlighted, and the rest are faded to the
background. Hence, time series meeting the constraints imposed by the brushes
are focused, while the remainder are de-emphasized, though they remain present
as context. This method can be interpreted programmatically as the visual analog
of a database query, or probabilistically as the conditional distribution for
the full series, given it passes through certain bounds.

In DOI trees, the viewer's attention is focused on a collection of high-interest
nodes, while the remaining lower-interest nodes are left on the fringes as
context. The implementation is modularized into two tasks -- the determination
of a DOI distribution over nodes in the tree and visual layout of a tree given
DOI assignments. The DOI distribution used in \citep{heer2004doitrees} places
maximal interest on the node that the user had most recently clicked, along with
all ancestors. The DOI for all other nodes is defined as the distance to the
closest maximal interest node. The layout step then trims low-interest subtrees
until the remaining nodes fit within a given screen size. By adjusting the
minimal DOI below which nodes are hidden, the user can transition between
node-specific and full-tree scales.

In linking, alternative representations of the same samples are placed
side-by-side in order to display covariation across views. A canonical
application is to linked scatterplot brushing \citep{becker1987brushing}. Here,
a scatterplot matrix gives the relationship between all pairs of variables.
~Points brushed in one scatterplot are then highlighted in all others. For
example, this helps the user determine whether an outlier in one dimension is an
outlier in others. Another instance of this idea links the results of
dimensionality reduction methods to displays of the raw data, as implemented by
XGobi and Cranvas, for example \citep{xie2013cranvas, swayne1998xgobi}. As in
timeboxes, linking can be interpreted as database queries or conditional
probabilities: given a subset of the series after conditioning on the values for
one set of features, what are the values for a second set
\citep{buja1996interactive}?

Finally, unrelated to established visualization principles, we note that our
work is deliberately grounded in the R software ecosystem. This connection is
made using the htmlwidgets package \citep{vaidyanathan2014htmlwidgets}. Not only
does linking R with D3 make these visualization methods more broadly accessible,
we hope to facilitate exchange between data modeling and interactive
visualization. Moreover, our tools are intentionally limited in scope --
designed to facilitate this dialog for a specific class of problems, rather than
providing a toolbox for generic types of visualization design. We believe that
this narrow context within a broad ecosystem strikes a balance between
problem-specificity and ease-of-use.

\subsection{Specific Proposals}\label{specific-proposals}

Our first proposed visualization technique is a minor modification of the DOI
tree. The standard DOI tree definition does not have any notion of data defined
at nodes, it is only used a device for navigating tree structures. A trivial
extension can encode scalar data at nodes: have the node radius reflect the
associated scalar value. To reinforce this effect, we can adjust the width of
the parent edge. When parent nodes have values equal to the sum of their
children, this creates the effect of values ``flowing'' from the root to leaves.
To help viewers make use of their domain knowledge, we have included a search
box that highlights paths to nodes with matching terms. Edges are ordered from
widest on the left to narrowest on the right. While this method can only
represent a single scalar-value per tree node, it suggests an approach to the
tree-structured differential abundance problem, which we call the DOI sankey.

In the DOI sankey, we split each edge in the DOI Tree across different groups.
For example, suppose we have the average counts for treatment and control groups
at each tree node. Every edge in the tree is split into two colors\footnote{We
  use the colorbrewer palette to facilitate readability
  \citep{brewer2003colorbrewer}}, with relative widths of the different colors
reflecting differences in sizes for the two groups. The overall width of each
edge represents the sum of values across all groups.

This display is designed to facilitate investigation of the tree structured
differential abundance question. For example, for a single node and a single
group, first compute the average abundance at that node among all samples in
that group. This will give the width for that group's color on the edge leading
to the specified node. Differentially abundant subtrees then correspond to
subtrees where some colors occupy more space than others. That is, this
representation makes it easier to identify points where the ``flows'' for
different groups diverge -- the colors begin to separate. The DOI principle
assists navigation across the tree structure, allowing focus on individual flow
structures without losing broader tree context.

Our third display is directed at the bacterial dynamics question. Here, two
panels are arranged one over the other; one displays all time series, while the
other displays all tree nodes, with node sizes reflecting the value at that node
averaged across all time points. For this reason, we call the display, timebox
trees. In the time series panel, we have directly implemented the timeboxes
idea. We then link the panels: when a set of series is highlighted by the
timeboxes, the associated tree nodes are also highlighted. For example,
timeboxes can be used to focus on a set of series with specific shape --
increased abundance after an ecological shock, for example -- and identify along
what subtrees this pattern is present. To further focus on specific elements, a
pan-zoom scented widget is provided \citep{willett2007scented}. The widget is a
miniature version of the full time series panel, equipped with a single brush
whose extent specifies the limits in the main time series panel. As in the DOI
trees and sankeys, a search bar can be used to highlight those series of
interest a priori.

The final display currently implemented in the package is the natural converse
of the timebox trees display. Rather than defining visual queries in terms of
time series, it defines queries using nodes in the tree. For this reason, we
call the display treeboxes. Rather than focusing on the intersection of brushes,
as in timebox trees, we focus on the union of brushed over nodes. This allows us
to highlight series associated with nodes on distant subtrees. This display is
also suited for the bacterial dynamics problem. For example, by highlighting all
nodes at one taxonomic level in the tree, we can easily summarize the time
series pattern for all the taxa at that level. Alternatively, focusing on all
the children below a single node makes it possible to see how much correlation
and competition there is between taxonomically similar bacteria. As in the
timebox trees display, a search box and pan-zoom scented widget are provided.

To be practically useful, the resulting visualizations must respond fluidly to
user interaction. As the data increase in scale, this fluidity can deteriorate
for two reasons. First, rendering many SVG elements in a framework like D3 is
costly\citep{johnson2008scalability}. While it is possible to use alternatives
-- HTML5's Canvas, for example -- it is often more challenging to implement
complex interactive behavior through them. Second, some of the dynamic queries
require a search over a many elements. These limitations are most pronounced in
the timebox tree display, which must search through all timepoints among all
time series whenever the brush is moved. The first, but not the second, concern
applies to treeboxes, while neither applies to DOI trees. Nonetheless, we feel
comfortable recommending timebox trees for data on the order of 500 tree tips
and 50 timepoints We note the sizes of the data sets in each case study below,
which each render fluidly, with the possible exception of the California housing
data, where there is a noticeable lag in the timebox tree and treebox displays.
In problems of larger size, we recommend a preliminary filtering or aggregation
step across nodes or, if the time series is smooth, across neighboring
timepoints, in order to avoid these potential scaling difficulties.

\section{Case Studies}\label{case-studies}

We now delve into applications on real data. Our goals are to illustrate
potential workflows that incorporate treelapse, describe the formulation of
questions that can be naturally investigated with our methods, and provide
example interpretations on treelapse output. Our examples are also chosen to
reflect the range of problem domains to which the package can be applied --
though it was motivated by applications to the microbiome, it is not tied to it.
More importantly, we argue that the visualization principles reviewed above can
substantively improve the practice of data analysis in the class of problems to
which we have limited ourselves.

\begin{table}
\centering
\begin{tabular}{lll}
Data            & Number of timepoints & Number of nodes \\
Antibiotics     & 56                 & 386             \\
Preterm births  & 216                & 318             \\
Housing prices  & 254                & 944             \\
Bikesharing     & 24                 & 819             \\
Global patterns & 500                & 51             
\end{tabular}
\caption{Problem dimensions for each of the case studies. For problems with
  dimensions larger than that in the housing prices problem, we recommend an
  initial summarization or filtering step to prevent performance
  issues. \label{problem-scaling}}
\end{table}

\subsection{Bacterial Dynamics of Antibiotics Time
Courses}\label{bacterial-dynamics-of-antibiotics-time-courses}

\citet{dethlefsen2008pervasive} investigated the effect of antibiotics on
bacterial community composition from an ecological perspective. The study tracks
the microbiome of three patients across ten months, with two five-day antibiotic
time courses separated by 6 months. Discerning the variation in resilience
across bacteria is important, considering the the role of bacteria in health and
not just disease.

We approach the data using the linked time and treebox views, after first
filtering low variance taxa and taking an \(\text{asinh}\) transformation. An
initial view, Figure \ref{fig:antibioticoverview}, reveals two dramatic drops in
the overall bacterial abundance time series during the antibiotics time courses.
Two more subtle effects are also suggested from this view,

\begin{itemize}
\item
  The second antibiotic treatment seems to have a more lasting effect,
  as the series take longer to return to their original values.
\item
  Some high level taxa appear relatively unaffected by the first
  antibiotic treatment. By more closely inspecting the display, we are
  able to identify these as members of the Bacteroidetes phylum, see
  Figure \ref{fig:antibioticbacteroidetes}.
\end{itemize}

\begin{figure}

{\centering \includegraphics[width=350px]{figure/annotated_antibiotic_overview}

}

\caption{The two antibiotic time courses are readily apparent even when only
  highlighting the most abundant taxa. We have annotated the figure with its
  three main components.}\label{fig:antibioticoverview}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=350px]{figure/antibiotic_bacteroidetes}

}

\caption{Introducing a second box into the timebox display identifies the
  Bacteroidaceae as a taxon only midly impacted by
  antibiotics.}\label{fig:antibioticbacteroidetes}
\end{figure}

Next, using the scented widget, we focus on the window around the second
antibiotic treatment. We apply the treebox display to compare then behavior of
different families of Firmicutes, Lachnospiraceae and Ruminoccocus. We suspect
that these taxa are associated with the delayed recovery after the second time
course. To investigate this, we input these family names in the search box to
isolate their positions on the tree; then we apply brushes to highlight the
series that contribute to these higher-level families. The resulting view is
given by Figure \ref{fig:antibioticfirmicutes}

\begin{figure}
{\centering \includegraphics[width=350px]{figure/antibiotic_firmicutes}

}

\caption{Zooming into the second antibiotic timecourse and highlighting the
  Lachnospiraceae and Ruminococcus, we see that the Ruminoccocus took more time
  to recover to pre-treatment levels.}\label{fig:antibioticfirmicutes}
\end{figure}

Alternatively, we can summarize each node by the average across its
descendants -- this brings attention to individual bacteria that may be
underlying some of the broader taxonomic patterns we have noted when
studying the subtree sums. For example, in Figure
\ref{fig:ruminococcus}, we highlight all families below order
Ruminococcus, suggesting that the decrease due to antibiotics occurs
uniformly across almost all families. A point that was not evident in
the earlier sum-across-descendants view is that, after the second
treatment of antibiotics, a few of the Ruminoccocus families recover
more rapidly than the rest, for example the Unc095d3 (highlighted in
red) are only briefly affected. In contrast, most families seem to
recover in unison after the first treatment.

\begin{figure}

{\centering \includegraphics[width=350px]{figure/ruminococcus}

}

\caption{By hovering over the Ruminoccocus branches, we see that there is a
  prolonged effect of the antibiotics time courses more or less uniformly across
  the lower taxonomic orders.}\label{fig:ruminococcus}
\end{figure}

Further, note that in this subtree averages view, the tree display has changed.
This is because, at each branching point, we place the node with larger average
value on the left. Figure \ref{fig:verrucomicrobiae} notes that the nodes at
the far left in the tree are associated with phylum Verrucomicrobiae,
corresponding to a large average abundance across time points. This phylum had
been previously obfuscated -- because there are not many leaves associated with
this phylum, the sum was small. Interestingly, the abundance of these bacteria
seems to \emph{increase} after the first antibiotics treatment. Be cautious,
however, that the average over only a few Verrucomicrobiae species will be a
more variable estimate than the averages over the more prevalent phyla.

\begin{figure}

{\centering \includegraphics[width=350px]{figure/verrucomicrobiae}

}

\caption{The subtree averages aggregation brings attention to the
  Verrucomicrobiae, which though only present as a few species, are each rather
  abundant. In particular, they seem to increase after the first antibiotic time
  course.}\label{fig:verrucomicrobiae}
\end{figure}

\subsection{Differential Bacterial Abundance and Preterm
Births}\label{differential-bacterial-abundance-and-preterm-births}

\citet{digiulio2015temporal}
tracked the abundance of bacteria in the vaginal microbiome during
pregnancy in an effort to study relationships between bacterial
community composition and preterm birth. Ideally, it would be possible
to develop clear bacterial signatures associated with preterm births.

Unlike the antibiotics study, we have measurements across more
individuals than we could reasonably inspect one at a time. While we
could average across all individuals, we will take our cue from
\citep{digiulio2015temporal}
place each sample into one of five Community State Types (CSTs),
identified via k-medoids. In that study, a linear model identified one
of these CSTs (CST 4) as significantly more diverse, further it appeared
associated with preterm births. Here, we corroborate this finding using
exploratory views.

Therefore, our focus here is on the differential abundance question,
rather than dynamics. We would like to provide visual representations of
differential abundance across CSTs and also between preterm and non-preterm
births. \citet{digiulio2015temporal} interpreted the CSTs using a heatmap, with
bacteria ordered according to a hierarchical clustering. By using the DOI sankey
instead, we can interpret the CSTs in their taxonomic context and at multiple
scales of taxonomic resolution. Further, while
\citet{digiulio2015temporal} focused on identifying associations between
preterm births and CSTs -- presumably because testing individual bacteria loses
power -- we can compare bacterial abundances between preterm and non-preterm
samples along subtrees, without requiring CSTs as an intermediary.

In Figure \ref{fig:pretermcsts}, we compare the 5 CSTs according to
their values along the subtree. Specifically, we took the average of all
samples within each CST to define values at the (species-level) leaf
nodes, and then aggregated the averages up to the root. It is
immediately clear that samples from CST 4 have much more taxonomic
diversity. Further, focusing on the Lactobacillaceae family, we note
that the differential abundance of these bacteria distinguishes the
remaining CSTs, see Figure \ref{fig:pretermcstslacto}.

Alternatively, in Figure \ref{fig:pretermpreterm}, we avoid working with CSTs,
displaying instead averages among samples associated with either preterm or term
births. The green and yellow edges are associated with preterm births -- we see
that they contribute more weight to phyla outside the Firmicutes. This is
consistent with the claim that CST 4, the most diverse of the CSTs, is
associated with preterm births.

\begin{figure}

{\centering \includegraphics[width=350px]{figure/preterm_csts}

}

\caption{The increased diversity among samples in CST 4 is represented by the
  relatively larger contribution of red edges to branches outside of the
  Firmicutes.}\label{fig:pretermcsts}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=350px]{figure/preterm_csts_lacto}

}

\caption{Zooming into the Lactobacillaceae family, we notice that the difference
  between the remaining four CSTs is related to which types of Lactobacillus are
  most prominent.}\label{fig:pretermcstslacto}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=300px]{figure/preterm_preterm}

}

\caption{Samples with high levels of phyla other than Firmicutes appear to be
  related to preterm births.}\label{fig:pretermpreterm}
\end{figure}

\subsection{Dynamics in Housing Prices}\label{zillow-study}

We next consider an application unrelated to the microbiome, but with relatively
clear hierarchical structure. Our data are downloaded from
Zillow\footnote{\url{http://www.zillow.com/research/data/}}, and give the Zillow
Home Value Indexes at the neighborhood level, across the country, computed
monthly between 1996 and 2016. In our display, we have taken the base 10 log of
these indexes. As our hierarchical structure, we use each neighborhood's
assignment to state, regional, county, and city levels. We represent each of
these coarser spatial categories using the average of all neighborhoods
contained in them. We have filtered down to the 890 neighborhoods in California;
rendering more neighborhoods while keeping all 246 timepoints causes the
interface to lag\footnote{See below for potential optimizations, however.}.

Our basic analysis revolves around geographic and temporal variation in home
prices. We are especially interested in the effect of the 2008 recession and any
variation in the lead-up to or recovery from this event. These questions can be
naturally framed using timebox trees and treeboxes.

We first study variation in prices at the city level. These series can be
highlighted using a single brush in treeboxes, since each spatial level is
displayed at the same height in the tree. The resulting display is given in
Figure \ref{fig:zillowcities}. From a high level, the housing bubble, decrease
in prices due to the recession, and subsequent recovery are readily apparent.
The city-level view makes it clear that not all neighborhoods were equally
affected by the recession -- richer cities plateaued at their 2008 prices,
middle-income cities saw moderate decreases, and poorer cities saw the most
significant declines in home prices. These more significant declines tended to
occur in cities that had recently seen rapid increases during the housing
bubble. Finally, we note that the range in home prices seems to have increased
after the recession, speaking to the differential long-term effect of the
recession on prices.

Alternatively, we can study the trajectories of home prices among neighborhoods,
conditional on their being middle-income before the recession. We generate the
sequence of views in Figures \ref{fig:zillowmiddlepre},
\ref{fig:zillowmiddleup}, and \ref{fig:zillowmiddledown} to this end. The first
of these figures isolates neighborhoods with middle incomes before the
recession, using a single timebox. Since there appears to be a divergence in
trajectories after the recession, we introduce a second post-recession timebox,
dragging it over series with higher and lower incomes during this second time
period. This is the content of Figures \ref{fig:zillowmiddleup} and
\ref{fig:zillowmiddledown}. Inspecting the highlighted tree nodes associated
with these series, we find that most of the middle-income series that increased
after the recession are associated with middle-income neighborhoods within the
coastal Southern California counties. In constrast, those middle income series
that saw decreases were mostly located in Central California and Oakland.

The previous analysis highlights the fact that, within even narrow geographic
regions, there can be substantial variation in prices. We can study this
directly using treeboxes. In Figure \ref{fig:zillowsf} we have highlighted all
series in San Francisco County. We see that, in 2016, prices range from around
\(10^{5.6} \approx \$400,000\) to \(10^{6.7} \approx \$5 \text{ million}\). So,
while all these neighborhoods tend to be among the more expensive ones in
California, prices can vary in a non-smooth way across geographic space.

We conclude this example with a caveat that the Zillow data are not
representative of all neighborhoods in California, only those with enough
listings on the site, so should be supplemented by other data sources for any
substantial decision-making.

\begin{figure}

{\centering \includegraphics[width=350px]{figure/zillow_cities}

}

\caption{California home prices at the city level, between 1996 and 2016. The
  effect of the 2008 recession is clear, and we have hovered over the San
  Bernardino series, to display the identity of one of the cities most strongly
  affected by the recession.}\label{fig:zillowcities}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=350px]{figure/zillow_middle_pre}

}

\caption{Neighborhoods with mid-range home prices before the recession are
  selected here. Note that the collection of series seems to widen after 2008 --
  we are interested in whether there are reliable predictors of these alternate
  trajectories, given their similar starting points.}\label{fig:zillowmiddlepre}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=350px]{figure/zillow_middle_up}

}

\caption{Among those neighborhoods with mid-range prices before the recession,
  we have selected those that recovered more rapidly. These appear to be located
  mainly in Los Angeles and San Diego counties.}\label{fig:zillowmiddleup}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=350px]{figure/zillow_middle_down}

}

\caption{Rather than selecting mid-range series that recovered quickly, we can
  isolate those whose prices remained depressed after the recession. These seem
  mostly to be located in Central California and the East San Francisco Bay
  Area.}\label{fig:zillowmiddledown}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=350px]{figure/zillow_sf}

}

\caption{To study the range in home prices within San Francisco County, we can
  select the associated nodes using a treebox.}\label{fig:zillowsf}
\end{figure}

\subsection{Sources of Variation in Bikesharing
Demand}\label{bikesharing-study}

Our next example is a study in bikesharing demand, included as an
example of analyzing collections of time series when there is no obvious
hierarchical structure a priori. The data are available at the UCI
Repository\footnote{\url{https://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset}}
and were originally collected by a Washington D.C.-based bikesharing
system for use in a Kaggle prediction competition. The data are hourly
measurements of bike demand, aggregated across all bikesharing stations,
over two years, along with supplemental weather data. In the
competition, participants were asked to predict the hourly demand on a
held-out test set. Here, we adopt a descriptive view instead, attempting
to characterize factors associated with variation in bikesharing demand.

Like the Zillow home prices application, we study this problem as one of
describing a large collection of related time series. Here, we consider
the demand during a single day to be one time series; this is a natural
choice considering the daily periodicity of bike demand. To arrange
these daily series along an interpretable tree structure, we apply a
regression tree relating the supplemental data to the bikesharing demand
\citep{breiman1984classification}. In more
detail, we built this tree by noting the ``two table'' structure of this
problem: one describes bike demand, the other holds the supplemental
data. In both, the rows index days, while the columns index either hours
or supplemental features. Our tree is the trained regression tree after
predicting demand at 8AM based on the supplemental data. We choose this
response because (1) we need a univariate response in order to apply
regression trees and (2) the more straightforwards reduction to
daily-average-demand fails to distinguish between weekdays and weekends,
whose series appear qualitatively very different from each other.

Given this response, the first split in the regression tree is
(unsurprisingly) the difference between weekends and weekdays. This is
emphasized in Figures \ref{fig:working} and \ref{fig:weekend}, respectively;
using timeboxes to isolate the two types of series highlight the left and right
sides of the tree, respectively. For a more subtle effect, we select the
internal nodes associated with the first split below the weekday vs. weekend
split; these are given in Figures \ref{fig:weekday2011} and
\ref{fig:weekday2012}. This suggests that weekday demand increased during the
second year.

\begin{figure}

{\centering \includegraphics[width=350px]{figure/working}

}

\caption{The two peaks at rush hour distinguish weekday series from the
  rest.}\label{fig:working}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=350px]{figure/weekend}

}

\caption{Unlike weekday demand, weekend demand is unimodal. The few weekday
  series with unimodal series seem to be associated with holidays. This is the
  case for New Years' Eve, which is currently hovered over in the
  tree.}\label{fig:weekend}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=350px]{figure/weekday_2011}

}

\caption{Weeday demand appears larger in 2012 than 2011 -- compare with the next
  figure.}\label{fig:weekday2011}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=350px]{figure/weekday_2012}

}

\caption{Weekday demand increased in 2012 -- compare with the previous
  figure.}\label{fig:weekday2012}
\end{figure}

In contrast to these general questions about daily demand, we could ask
a more granular question about specific time windows. For example, what
characterizes days on which there is larger than average demand after
midnight? We can select these series after first zooming into this time window.
Figure \ref{fig:warmweekend} reveals that the highlighted series are associated
with the warm-weekend split, which seems quite reasonable in retrospect.

\begin{figure}

{\centering \includegraphics[width=350px]{figure/warm_weekend}

}

\caption{The samples with highest night demand tend to fall on warm
  weekends.}\label{fig:warmweekend}
\end{figure}

Finally, we can study the behavior of the regression tree itself using
the DOI sankey. Here, we group samples according to their quintile of
8AM demand and then count the abundance of the groups flowing down
different branches. We find that the quintiles are each rather strongly
separated after descending even a few steps down the regression tree --
for example, Figures \ref{fig:weekday2011} and \ref{fig:weekday2012}
focus on 2011 vs. 2012 split among weekday samples, showing that this
split distinguishes between samples falling in the second and third
quintiles of 8AM demand.

\begin{figure}

{\centering \includegraphics[width=350px]{figure/bike_sankey}

}

\caption{We can interpret the regression tree using an interactive DOI
  representation.}\label{fig:bikesankey}
\end{figure}

This interactive representation of regression trees is potentially more
useful on larger trees that cannot be easily parsed in a single view; in
this sense the bikesharing tree is relatively simple. In our ideal data
analysis workflow, we imagine the analyst applying interactive
visualization and modeling techniques in an iterative, nonlinear
fashion.

\subsection{Hierarchically Clustering the Global Patterns Data}\label{global_patterns}

Each of the timebox tree and treebox examples presented so far have
focused on data with a clear time component. We note however that these
techniques could alternatively be applied to high-dimensional data, via
the use of parallel coordinates \citep{inselberg1991parallel}. The usual
parallel coordinates challenges remain, namely selecting scales for and an
ordering across the different coordinates, but the linking and
focus-plus-context ideas can still be employed in this setting. Here we
provide an implementation of this idea on a dataset comparing microbiomes across
various ecological environments \citep{caporaso2011global}, which is publicly
accessible through the phyloseq R package \citep{mcmurdie2013phyloseq}.

The original Global Patterns data consists of 26 samples across 9
environments (for example, freshwater and soil). In each site, there are
counts across 19216 taxa -- to simplify visualization, we filter to the
500 most abundant taxa.

We hierarchically cluster these 26 samples based on the 500 most
abundant taxa, using complete linkage on the UniFrac distance. Figure
\ref{fig:gptimebox} displays the resulting hierarchy together with a
parallel coordinates view of the \(\text{asinh}\) transformed taxa.

\begin{figure}
  {
    \centering
    \includegraphics[width=225px]{figure/gp_cluster1}
    \includegraphics[width=225px]{figure/gp_cluster2}
}

\caption{An application to the Global Patterns demonstrates how linking in
  treelapse can be applied to combine hierarchical clustering and parallel
  coordinates views.}\label{fig:gptimebox}
\end{figure}

In Figure \ref{fig:gptimebox}, we compare two subclusters from the
hierarchical clustering tree, after zooming to a few of the bacteria
that distinguish between the clusters. Upon revisiting the original
data, it becomes clear that the samples highlighted on the left come
from freshwater samples, while those on the right come from soil and
skin, and looking up taxonomic groups associated with the distinguishing
bacteria confirms this. For example, many of the species with high
abundances in the left figure come from order Oceanospirillales.

\subsection{Inspecting Confirmatory Analysis}\label{structssi}

In addition to facilitating exploratory study, treelapse has potential
value as a device for inspecting confirmatory analysis. We provide an
illustration extending an example from \citep{callahan2016bioconductor}, which
formally tested bacterial species for association with age in a sample of mice.
The testing approach advocated there is particularly well-suited to
visualization with treelapse, since it sought to detect associations at multiple
levels of phylogenetic resolution, using statistical tools developed in
\citep{yekutieli2008hierarchical, sankaran2014structssi}.

The data of interest in \citep{callahan2016bioconductor} are bacterial counts
collected across old and young mice. After variance-stabilizing these counts
using DESeq2 \citep{love2014moderated}, a $t$-test was applied to each node in a
phylogenetic tree, comparing abundances between old and young mice. To account
for multiple testing, we employ the structSSI algorithm
\citep{yekutieli2008hierarchical, sankaran2014structssi} along with methods
available in the \texttt{multtest} package \citep{pollard2005multiple}.

To interpret the results, we apply timebox trees. Our goals are to (1) identify
subtrees with consistently elevated differential abundance across age groups and
(2) compare alternative multiple testing adjustment procedures. Our approach is
to display the negative-log raw and adjusted $p$-values for each node, with
alternative methods compared via parallel coordinates. One view of the resulting
display is captured in Figure \ref{fig:structssi}. First, we see that
significant nodes tend to be significant across all methods -- the ordering
between different series appears stable. Interestingly, the Sidak one-step and
structSSI procedures seem to have lower power than the others, including
conservative FWER-controlling methods, like the Bonferroni procedure. Further,
in this application, FDR-controlling techniques do not seem to offer notably
different adjusted $p$-values, relative to those controlling FWER. This suggests
that, for this problem, bacteria are either strongly associated with age, or not
associated at all, so that there is little gain from using more sensitive
procedures.

\begin{figure}

{\centering \includegraphics[width=250px]{figure/structssi}

}

\caption{Viewing a tree of $p$-values across different methods highlights two
  subtrees with strong associations with mouse age, across several testing
  procedures.}\label{fig:structssi}
\end{figure}

Further, selecting series with strong association between abundance and age, two
major subtrees are brought to the forefront. Separately querying the the
taxonomic identities of these bacteria reveals that they are two subgroups of
Clostridia, which is consistent with the analysis of
\citep{callahan2016bioconductor}. More than this specific analysis outcome, this
view demonstrates that interactive visual inspection of results from
confirmatory analysis provides deeper insight than the standard practice of
printing tables of (adjusted or unadjusted) $p$-values: the relationship between
significant nodes is only clear upon visualization on the tree.

\section{Conclusion and Future Work}\label{conclusion}

Here, we have reviewed some fundamental principles of data visualization and
described their implementation in a new treelapse package. Further, we have
provided examples of the practical usefulness of these principles in real-world
data analysis situations.

This package has only developed basic ideas, and there are a number of
potentially useful extensions worth exploring. For example, we have not
considered the principle of arrangement in our visualizations
\citep{buja1996interactive}, though many of our conclusions were based on
comparing alternative selections of the same display. We could imagine faceting
our displays across groups to make these types of comparisons more accessible.
Further, we have only worked with the DOI distribution described in
\citep{heer2004doitrees}. It would be interesting to define a more statistical
notion of interest along nodes, based on cognostics, for example
\citep{hafen2013trelliscope, friedman2002john}. A simple extension could be to
allow graph layouts instead of trees in time and treebox displays, for data that
cannot be coerced into a hierarchical structure. Finally, if these ideas turn
out to be useful in practice, it would be valuable to modularize the basic
visualization layouts and relationships into a library, allowing the wider
community to construct novel linked, interactive graphics with minimal effort.

In summary, we have built an easily accessible R package for visualization
techniques in a very specific methodology problem -- analysis of differential
abundance and dynamics in hierarchically structured data -- that appears in a
variety of application domains. We have leveraged a link between R and D3
\citep{vaidyanathan2014htmlwidgets} to create visualizations during the
exploratory phase of data analysis; in this way our work is a departure from the
culture of polished, journalistic visualizations prioritized by the D3 community
and is more closely aligned with the vision in \citep{de2003visual} of more
tightly integrating data visualization and statistical analysis techniques.
Finally, we have given a series of examples to demonstrate how the general
visualization techniques of focus-plus-context and linked brushing can be
practically incorporated into a range of practical analysis workflows, from
studying the impact of bacteria on human health to better allocating units in
commuter bikesharing systems.

\bibliographystyle{plainnat}
\bibliography{bibliography}

\end{document}