### 2017-09-23 supplement.tex

\documentclass{article}
\usepackage{url}

\title{Supplemental Materials}
\author{Kris Sankaran}

\begin{document}
\maketitle

\section{Supplemental Materials}

\subsection{Links}
\label{sec:supp_links}

\begin{itemize}
\item treelapse home page: \url{https://krisrs1128.github.io/treelapse/}
\item treelapse development page: \url{https://github.com/krisrs1128/treelapse}
\item treelapse video demo: \url{https://www.youtube.com/watch?v=EcmYBRMVMbI&t=14s}
\item Home prices data: \url{http://www.zillow.com/research/data/}
\item Bikesharing data: \url{https://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset}
\end{itemize}

\end{document}